---
output:
  html_document: default
  pdf_document: default
---
Loan Data Case Study by Timothy Short
========================================================
```{r echo=FALSE}
knitr::opts_chunk$set(warning=FALSE, message=FALSE)
```


```{r echo=FALSE, packages}
library(ggplot2)
library(dplyr)
library(gridExtra)
```

```{r echo=FALSE, Load_the_Data}
df = read.csv('data/prosperLoanData.csv', header=T)
```

There are 113,937 observations with 81 variables to analyze. Variables include
basic loan information and applicant information, to detailed description of
the loan backer and financial institution information. Most variables are
numerica with String variables as factor variables.

There are many different case studies to analyze, from trends of lower income
to upper income customers, to analyzing what products customers will seek
to finance, to the credit worthiness of the borrower and / or the loan.

### Objective
My goal is to consider the loan status and see which variables seem to impact
and help identify trends among a borrower's loan status. Specifically, I am
interested to discover which factors help indicate and predict the riskiness
of a loan and the likelihood of the loan going into default or charged off.

Below are some initial data investigations. Factors include income range,
employment status, credit score of the borrower, purpose of the loan, loan
amount, and many other factors as explored below and throughout this first phase
of explording the data.

```{r echo=FALSE, eval=FALSE, InitialObservations}
head(df)
names(df)

print(summary(na.omit(df)))
#Borrower Profile
summary((df$CreditScoreRangeLower + df$CreditScoreRangeLower)/2)
table(df$OpenCreditLines)
summary(df$IncomeRange)
summary(df$DebtToIncomeRatio)
table(df$TotalInquiries)
table(df$EmploymentStatus)
table(df$IsBorrowerHomeowner)

Current Loan Data
count(df$LoanStatus)
summary(df$LoanCurrentDaysDelinquent)
table(df$CreditGrade)
summary(df$MonthlyLoanPayment)
```

Make preliminary changes to data set

```{r echo=FALSE, nitializeVariables}
#Calculate the mean credit score (upper - lower)
df$CreditScoreMean = (df$CreditScoreRangeUpper + df$CreditScoreRangeLower)/2

#Factor credit grade and income levels
df$CreditGrade = factor(df$CreditGrade,
  levels = c('NC', 'HR', 'E', 'D', 'C', 'B', 'A', 'AA'))

df$IncomeRange = factor(df$IncomeRange,
  levels = c('$0', '$1-24,999', '$25,000-49,999', '$50,000-74,999',
             '$75,000-99,999', '$100,000+',  'Not displayed', 'Not employed'))

#Define loan status category summary: then factor
  #Good = loan status of Current, Completed, or Final Payment
  #Late = loan status of payment pas due date
  #Bad = loan status of Cancelled, Charged Off, or Defauly
df$LoanStatusBinary =
  ifelse(df$LoanStatus=='Current' | df$LoanStatus=='Completed' |
           df$LoanStatus=='FinalPaymentInProgress', 'Good',
    ifelse(df$LoanStatus=='Cancelled' | df$LoanStatus=='Chargedoff' |
           df$LoanStatus=='Defaulted','Bad',
      'Late'))
df$LoanStatusBinary = factor(df$LoanStatusBinary,
      levels = c('Bad','Late','Good'))

#Income in annual terms
df$StatedAnnualIncome = df$StatedMonthlyIncome * 12

#Define variable analysis as array
borrowerProfile = c( "CreditScoreMean" , "OpenCreditLines" , "IncomeRange" ,
              "DebtToIncomeRatio" , "TotalInquiries" ,
              "EmploymentStatus" , "IsBorrowerHomeowner" )

loanData = c( "LoanStatus" , "LoanCurrentDaysDelinquent" ,
              "CreditGrade" , "MonthlyLoanPayment")

listingNames = c("Debt Consolidation", "Home Improvement", "Business",
  "Personal Loan", "Student Use", "Auto", "Other", "Baby&Adoption", "Boat",
  "Cosmetic Procedure", "Engagement Ring", "Green Loans", "Household Expenses",
  "Large Purchases", "Medical/Dental", "Motorcycle", "RV", "Taxes",
  "Vacation", "Wedding Loans")
```

## Univariate Plots

```{r echo=FALSE, 'Univariate_Plots: LoanData'}
#Bar graph to show the count of each loan status
ggplot(aes(x=LoanStatusBinary), y=count(LoanStatusBinary), data=df) +
  geom_bar() + 
  theme(axis.text.x = element_text(angle=45, vjust=.95, hjust=.95)) +
  ggtitle('Counts of Loan Status')
  #ggsave('visualizations/univariate/loanstatus.png')

#Bar graph to show listing category
ggplot(data=df, aes(x=ListingCategory..numeric.)) +
  geom_histogram(bins=20, binwidth = .5) +
  scale_x_continuous(limits = 
    c(0,21), breaks=seq(1,20,1), labels = listingNames) +
  theme(axis.text.x = element_text(angle=45, vjust=.95, hjust=.95)) +
  xlab('Listing Category') + 
  ggtitle('Loan Listing Category')
  #ggsave('visualizations/univariate/loanpurpose.png')

#Loan Amount
ggplot(df, aes(x=LoanOriginalAmount)) +
  geom_histogram() +
  ggtitle('Distribution of Loan Amounts')
  #ggsave('visualizations/univariate/loanamount.png')

#Loan Rates
ggplot(df, aes(x=BorrowerAPR)) +
  geom_histogram() +
  ggtitle('Distribution of Loan Rates')
  #ggsave('visualizations/univariate/loanrates.png')

#Histogram to show the counts of days for loan deliquency
ggplot(aes(x=LoanCurrentDaysDelinquent), data=subset(df, 
      df$LoanCurrentDaysDelinquent>0)) +
  geom_histogram(binwidth = 30) +
  scale_x_continuous(breaks=seq(1,365*30,365), labels=c(0:29)) +
  xlab('Year Delinquent') +
  ggtitle('Distribution of Loan Delinquency in Years')
  #ggsave('visualizations/univariate/daysdelinquent.png')

#Bar grapch to show counts of different credit grades of the loan
ggplot(aes(x=CreditGrade), y=count(CreditGrade), data=subset(df,
      df$CreditGrade!="")) +
  geom_bar() + 
  ggtitle('Distribution of Loan Credit Grades')
  #ggsave('visualizations/univariate/creditgrade.png')

#Boxplot to show distribution of monthly payments
ggplot(data=df, aes(x=MonthlyLoanPayment)) +
  scale_y_continuous(limits=c(0,1500)) +
  geom_histogram() +
  ggtitle('Distribution of Monthly Payment')
  #ggsave('visualizations/univariate/monthlypayment.png')
```


```{r echo=FALSE, 'Univariate_Plots: Borrower Profile'}
#Histogram to show total credit inquiries of each loan's borrower
ggplot(aes(x=TotalInquiries), y=count(df$TotalInquiries), data=df) +
  geom_histogram(binwidth = 1) +
  scale_x_continuous(limits=c(-1,25)) +
  ggtitle('Distribution of Total Credit Inquiries\nRange: 0 to 25')
  #ggsave('visualizations/univariate/totalinquiries.png')

#Histogram distribution of current open credit lines of each loan's borrower
ggplot(data=df, aes(x=OpenCreditLines)) +
  geom_histogram() +
  ggtitle("Distribution of Borrower's Open Credit Lines")
  #ggsave('visualizations/univariate/opencreditlines.png')

#Bar graph to show distribution of borrower's income range
ggplot(data=df, aes(x=IncomeRange)) +
  geom_bar() +
  theme(axis.text.x = element_text(angle=45, vjust=.95, hjust=.95)) +
  ggtitle('Distribution of Borrowers Income')
  #ggsave('visualizations/univariate/incomerange.png')

#Histogram to show distribution of borrowers mean credit score
ggplot(data=df, aes(x=CreditScoreMean)) +
  xlim(450, 850) +
  geom_histogram(binwidth=10) +
  ggtitle('Distribution of Borrowers Mean Credit Score')
  #ggsave('visualizations/univariate/creditscores.png')

#Histogram of borrowers debt-to-income ratio
p1 = ggplot(aes(x=DebtToIncomeRatio), data=df) +
  geom_histogram(bins=50) +
  ggtitle('Distribution of Borrowers Debt to Income Ratio')
p2 = ggplot(aes(x=DebtToIncomeRatio), data=df) +
  xlim(0,1) +
  geom_histogram() +
  ggtitle('Distribution of Borrowers Debt to Income Ratio\nFrom 0 to 1')
p3 = ggplot(aes(x=DebtToIncomeRatio), data=df) +
  xlim(1,10) +
  geom_histogram() +
  ggtitle('Distribution of Borrowers Debt to Income Ratio\nFrom 1 to 10')
grid.arrange(p1, p2, p3)
g = arrangeGrob(p1, p2, p3)
#ggsave('visualizations/univariate/debtincomeratio.png', g)

#Bar graph of borrowers employment status
ggplot(aes(x=EmploymentStatus), y=count(EmploymentStatus), data=subset(df,
      df$EmploymentStatus!="")) +
  geom_bar() +
  ggtitle('Distribution of Employment Status')
  #ggsave('visualizations/univariate/employmentstatus.png')

#Bar graph of borrowers homeownersip
ggplot(aes(x=IsBorrowerHomeowner), y=count(IsBorrowerHomeowner), data=df) +
  geom_bar() +
  ggtitle('Distribution of Homeowner Status')
  #ggsave('visualizations/univariate/homeowner.png')
```


```{r echo=FALSE, SummaryCalculations}
#Proportions of loan status
propGoodStatus = length(which(df$LoanStatusBinary == 'Good')) / nrow(df)
propBadStatus = length(which(df$LoanStatusBinary == 'Bad')) / nrow(df)
propLateStatus = length(which(df$LoanStatusBinary == 'Late')) / nrow(df)

#Proportion of income brackets
summary(df$IncomeRange)[2]/ nrow(df)

#Loan analysis stats:
table(df$ListingCategory..numeric.)[5]/ nrow(df)
summary(df$BorrowerAPR)
summary(df$LoanOriginalAmount)
table(df$CreditGrade)
summary(df$MonthlyLoanPayment)
table(df$Term)[2] / sum(table(df$Term))

#Borrower profile stats:
(table(df$EmploymentStatus)[2] + table(df$EmploymentStatus)[3] +
  table(df$EmploymentStatus)[7] + table(df$EmploymentStatus)[8]) / nrow(df)
table(df$IsBorrowerHomeowner)

#Credit analysis
summary(df$DebtToIncomeRatio)
summary(df$CreditScoreMean)
table(df$CreditGrade)
```


### Univariate Analysis

**Loan Status:** 83.2% of loans are in good standing, while 14.9% of loans are
in bad standing and 1.8% of loans are late.

**Income Distributions:** Borrowers income brackets follow a fairly normal
distirbution. Low and middle-income earners make up the bulk of the dataset,
with \$25,000-49,999 and \$50,000-74,999 making up 28% and 27% respectively.
Both upper income brackets of \$75,000-99,999 and \$100,000+ make up 15% each,
leaving about 8% of earnings making below \$25,000 or reporting no income.

**Loan Analysis:** Of the 20 different loan types, the most common loan purpose
is for debt consolidation, making up over 50% of all reported loans. Besides
Other (or uncategorized) loans, home improvements, businesses, personal loans
round are next largest, each taking about 3-7% of the loans. Loan rates follow a
slightly normal / slightly right skewed distribution, with a median rate or
20.9%. Loan amounts of very right skewed with a median of \$6,500 and mean of
\$8,337. The median monthly loan payment is \$218 with a 36-month term.

**Borrower's Profile:** Over 83% of borrowers had a kind of formal employment,
while the remaining 17% was either unemployed, or the information was either
ambiguous or unavailable. About half of borrowers were homeowners.

**Credit Analysis:** Borrowers have a median credit score of 689.5 and a very
right skewed distribution of debt-to-income ratio with median of .22. Both the
open credit lines and total credit inquiries are strongly right skewed also.
The loan's credit grade follows a somewhat normal distribution with C-grade.

**Next:** I am interested in investigating correlations to the status of the
loan. Specifically, which factors impact the likelihood of a loan resulting in
a bad status? How well have we anticipated potential bad loans? Which borrowers
and which factors help indicate if the loan will get paid off? What makes a
loan risky?

## Bivariate Plots

```{r echo=FALSE, Bivariate_Plots}
#Analyzing Loan Status
ggplot(data=df, aes(x=ListingCategory..numeric., y=LoanStatusBinary)) +
  geom_jitter(alpha=.1, size=.01, width=.3, height=.3) +
  scale_x_continuous(limits = c(1,21), breaks=seq(1,20,1), labels = listingNames) +
  theme(axis.text.x = element_text(angle=90, vjust=.5, hjust=.95)) +
  xlab('Loan Listing Type') +
  ggtitle('Loan Status by Loan Category')
  #ggsave('visualizations/bivariate/loanstatus_loantype.png')

ggplot(data=df, aes(x=BorrowerAPR, y=LoanStatusBinary)) +
  geom_jitter(alpha=.1, size=.1) +
  ggtitle('Loan Status by APR')
  #ggsave('visualizations/bivariate/loanstatus_apr.png')

ggplot(aes(x=EmploymentStatus, y=LoanStatusBinary),
       data=subset(df, df$EmploymentStatus!='Not available' &
        df$EmploymentStatus!='Other' & df$EmploymentStatus!='')) +
  geom_jitter(alpha=.02, size=.2) +
  theme(axis.text.x = element_text(angle=45, vjust=.95, hjust=.95)) +
  ggtitle('Loan Stats by Employment Status')
  #ggsave('visualizations/bivariate/loanstatus_employment.png')

ggplot(aes(x=IncomeRange, y=LoanStatusBinary),
       data=subset(df, df$EmploymentStatus!='Not available' &
        df$EmploymentStatus!='Other' & df$EmploymentStatus!='')) +
  geom_jitter(alpha=.02, size=.3) +
  theme(axis.text.x = element_text(angle=45, vjust=.95, hjust=.95)) +
  ggtitle('Loan Status by Income')
  #ggsave('visualizations/bivariate/loanstatus_income.png')

ggplot(aes(x=LoanOriginalAmount, y=LoanStatusBinary),
       data=subset(df, df$EmploymentStatus!='Not available' &
        df$EmploymentStatus!='Other' & df$EmploymentStatus!='')) +
  geom_jitter(alpha=.01, size=1, height=.3) +
  theme(axis.text.x = element_text(angle=45, vjust=.95, hjust=.95)) +
  ggtitle('Loan Status by Loan Amount')
  #ggsave('visualizations/bivariate/loanstatus_loanamount.png')

ggplot(aes(x=LoanStatusBinary), y=count(LoanStatusBinary), data=df) +
  geom_bar() +
  facet_wrap(~IsBorrowerHomeowner) +
  ggtitle('Loan Status by Homeowernship Status')
  #ggsave('visualizations/bivariate/loanstatus_homeownership.png')

#Analyzing the type of loan
ggplot(data=df, aes(x=ListingCategory..numeric., y=LoanOriginalAmount)) +
  stat_summary(fun.y='median', geom='bar') +
  scale_x_continuous(limits = c(0,21), breaks=seq(1,20,1), labels = listingNames) +
  theme(axis.text.x = element_text(angle=45, vjust=.95, hjust=.95)) +
  geom_hline(yintercept = 6500) +
  xlab('Loan Listing Type') +
  ylab('Median Original Loan Amount') +
  ggtitle('Median Loan by Loan Type')
  #ggsave('visualizations/bivariate/loanamount_loantype.png')

ggplot(aes(x=ListingCategory..numeric., y=CreditGrade), data=subset(df,
      df$CreditGrade!="")) +
  geom_jitter(alpha=.1, size=.3) +
  scale_x_continuous(limits = c(1,21), breaks=seq(1,20,1), labels = listingNames) +
  theme(axis.text.x = element_text(angle=45, vjust=.95, hjust=.95)) +
  geom_vline(xintercept=median(df$CreditScoreMean), size=.5) +
  xlab('Loan Listing Type') +
  ylab('Credit Grade') +
  ggtitle('Credit Grade by Loan Listing')
  #ggsave('visualizations/bivariate/creditgrade_loantype.png')

ggplot(data=df, aes(x=ListingCategory..numeric., y=BorrowerAPR)) +
  stat_summary(fun.y='median', geom='bar') +
  scale_x_continuous(limits = c(0,21), breaks=seq(1,20,1), labels = listingNames) +
  theme(axis.text.x = element_text(angle=45, vjust=.95, hjust=.95)) +
  xlab('Loan Listing Type') +
  ggtitle('Borrowers Median APR by Loan Type')
  #ggsave('visualizations/bivariate/apr_loantype.png')

ggplot(aes(x=IncomeRange, y=ListingCategory..numeric.), data=subset(df,
    df$IncomeRange!='Not employed' & df$IncomeRange!='Not displayed' & df$IncomeRange!='$0')) +
  geom_jitter(alpha=.02, size=.3) +
  scale_y_continuous(limits = c(1,21), breaks=seq(1,20,1), labels = listingNames) +
  theme(axis.text.y = element_text(vjust=.95, hjust=.95)) +
  ylab('Loan Listing Type') +
  ggtitle('Loan Type by Income')
  #ggsave('visualizations/bivariate/loantype_income.png')

#Analyzing the borrower
ggplot(aes(x=IncomeRange, y=CreditScoreMean), data=subset(df, df$CreditScoreMean > 50)) +
  geom_jitter(alpha=.03) +
  ggtitle('Mean Credit Scores by Income') +
  theme(axis.text.x = element_text(angle=45, vjust=.95, hjust=.95)) +
  ggtitle('Credit Score by Income Range')
  #ggsave('visualizations/bivariate/creditscore_income.png')

ggplot(aes(x=DebtToIncomeRatio, y=CreditScoreMean), data=df) +
  xlim(0,1) +
  geom_point(alpha=.05) +
  ggtitle('Credit Score by Debt-to-Income Ratio')
  #ggsave('visualizations/bivariate/creditscore_debttoincome.png')

ggplot(aes(x=EmploymentStatus, y=CreditScoreMean), data=df) +
  stat_summary(fun.y='median', geom='bar') +
  ggtitle('Credit Score by Employment Status')
  #ggsave('visualizations/bivariate/creditscore_employment.png')


#Analyzing Credit
ggplot(aes(x=CreditScoreMean, y=CreditGrade), data=subset(df,
    df$CreditGrade!="" & df$CreditGrade!='NC')) +
  geom_jitter(alpha=.1, size=.1) +
  xlim(400,900) +
  ggtitle('Loans Credit Grade by Credit Score')
  #ggsave('visualizations/bivariate/creditgrade_creditscore.png')

ggplot(aes(x=CreditScoreMean, y=BorrowerAPR), data=subset(df,
    df$CreditGrade!="" & df$CreditGrade!='NC')) +
  geom_jitter(alpha=.1, size=.5) +
  xlim(400,900) +
  ggtitle('APR  by Credit Score')
  #ggsave('visualizations/bivariate/apr_reditscore.png')

ggplot(aes(x=IncomeRange, y=BorrowerAPR), data=subset(df,
    df$CreditGrade!="" & df$CreditGrade!='NC')) +
  geom_jitter(alpha=.1, size=.5) +
  theme(axis.text.x = element_text(angle=45, vjust=.95, hjust=.95)) +
  ggtitle('APR  by Income')
  #ggsave('visualizations/bivariate/apr_income.png')

ggplot(aes(x=DebtToIncomeRatio, y=BorrowerAPR), data=subset(df,
    df$CreditGrade!="" & df$CreditGrade!='NC')) +
  geom_jitter(alpha=.1, size=.5) +
  xlim(0,1) +
  ggtitle('APR  by Debt-to-Income Ratio')
  #ggsave('visualizations/bivariate/apr_debttoincome.png')
```

```{r echo=FALSE, eval=FALSE, BivariateCalculations}
#Average APR by income bracket
summarize(group_by(df,IncomeRange),
          meanAPR = mean(BorrowerAPR, na.rm=T),
          meanCreditScore = mean(CreditScoreMean, na.rm=T),
          meanLoanAmount = mean(LoanOriginalAmount, na.rm=T)
      )
```


### Bivariate Analysis

After analyzing several varaiables agains loan status, borrower's interest rates,
and credit scores, we are able to draw several simple conclusions.

**Interest Rate:** Interest rates and loan status have an inverse relationship.
Loans with higher interest rates were more likely to be in bad standing as loans
with lower interest rates. This is not too surprising, as riskier loans have
higher interest rates and therefore cost more.

**Loan Amount:** Loan amounts and loan status tend to have a direct relationship.
Most of the loans in bad standing were less than $5,000, where as loan amounts
increased, it was more likely that they stayed current or were paid off. Loan
amounts also have a direct relationship with income brackets. Again, not too
surprising as higher income persons would presumably be less likely to borrow
smaller amounts of money.

**Loan Categories:** As previous mentioned, most loans were for the purpose
of debt consolidation. After analyzing the 20 different loan categories and
they cooresponding loan status, income brackets, interest rates, loan amounts,
I determined that further multivariate analysis needs to be conducted.

**Other Observations:**
- There seems to be a direct relationship with homeownership and loan status;
although homeowners represent about half of the demographic, more homeowners had
loans in better standing and fewer homeowners had loans in bad standing compared
with non-homeowners.

**APR Relationships:**
* Directly related to debt-to-income
* Inversely related to income
* Inversely related to credit score

**Credit Score Relationships:**
* Inverse relationship to APR (as mentioned above)
* Directly related to credit grade of loan
* Ambiguous relationship to debt to income (requires further analysis)
* Ambiguous relationship to employment (requires further analysis)
* Directly related to income

#### Anticipating multi-variate analysis

Bivariate analysis was able to show a few relationships between the loan type,
borrower's profile, and the loan status. However, we will need further analysis
to make better predictions and conclusions as to what factors are the true
indicators of a loan's status.

## Multivariate Plots

```{r echo=FALSE, Multivariate_Plots}
ggplot(aes(x=StatedAnnualIncome, y=LoanOriginalAmount), data=subset(df,
    df$IncomeRange!='Not employed' & df$IncomeRange!='Not displayed')) +
  geom_jitter(aes(color=LoanStatusBinary), size=.2) +
  scale_x_continuous(limits = c(1,250000)) +
  theme(axis.text.x = element_text(angle=90, vjust=.5, hjust=.95)) +
  ggtitle('Loan Amount by Income and Status')
  #ggsave('visualizations/multivariate/loanamount_income-status.png')

ggplot(aes(x=ListingCategory..numeric., y=StatedAnnualIncome), data=subset(df,
    df$IncomeRange!='Not employed' & df$IncomeRange!='Not displayed')) +
  geom_point(aes(color=LoanStatusBinary), size=.2) +
  scale_y_continuous(limits = c(1,250000)) +
  scale_x_continuous(limits = c(1,21), breaks=seq(1,20,1), labels = listingNames) +
  theme(axis.text.x = element_text(angle=90, vjust=.5, hjust=.95)) +
  xlab('Loan Type') +
  ggtitle('Income by Listing Category by Loan Status')
  #ggsave('visualizations/multivariate/income_category-status.png')

ggplot(aes(x=ListingCategory..numeric., y=LoanOriginalAmount), data=subset(df,
    df$IncomeRange!='$0' &  df$IncomeRange!='Not employed' &
      df$IncomeRange!='Not displayed')) +
  geom_point(aes(color=IncomeRange), alpha=.7, size=.5) +
  scale_color_brewer(palette='Blues') +
  scale_x_continuous(limits = c(1,21), breaks=seq(1,20,1), labels = listingNames) +
  theme(axis.text.x = element_text(angle=90, vjust=.5, hjust=.95, size=6)) +
  facet_wrap(~LoanStatusBinary) +
  ggtitle('Loan Amounts by Listing Category, by Income and Loan Status')
  #ggsave('visualizations/multivariate/loanamount_loantype-income-status.png')

ggplot(aes(x=LoanOriginalAmount, y=BorrowerAPR), data=subset(df,
    df$IncomeRange!='$0' &  df$IncomeRange!='Not employed' &
      df$IncomeRange!='Not displayed')) +
  geom_point(aes(color=LoanStatusBinary), size=.2) +
  facet_wrap(~IsBorrowerHomeowner) +
  ggtitle('APR by Loan Amount and Homeowner Status')
  #ggsave('visualizations/multivariate/apr_loanamount-status-homeowner.png')

ggplot(aes(x=LoanOriginalAmount, y=BorrowerAPR), data=subset(df,
    df$IncomeRange!='$0' &  df$IncomeRange!='Not employed' &
      df$IncomeRange!='Not displayed')) +
  geom_jitter(aes(color=IncomeRange), alpha=.3, size=.5) +
  theme(axis.text.x = element_text(angle=45, vjust=.95, hjust=.95)) +
  scale_color_brewer(palette='Blues') +
  facet_wrap(~LoanStatusBinary) +
  ggtitle('APR by by Loan Amount\nGrouped By Income and Faceted by Status')
  #ggsave('visualizations/multivariate/apr_amount-income-status.png')
  
ggplot(aes(x=CreditScoreMean, y=BorrowerAPR), data=subset(df,
    df$IncomeRange!='$0' &  df$IncomeRange!='Not employed' &
      df$IncomeRange!='Not displayed')) +
  geom_jitter(aes(color=LoanStatusBinary), size=.2) +
  theme(axis.text.x = element_text(angle=45, vjust=.95, hjust=.95)) +
  ggtitle('APR by Credit Score and Status')
  #ggsave('visualizations/multivariate/apr_creditscore-status.png')
  
ggplot(aes(x=IncomeRange, y=LoanOriginalAmount), data=subset(df,
    df$IncomeRange!='$0' &  df$IncomeRange!='Not employed' &
      df$IncomeRange!='Not displayed')) +
  geom_jitter(aes(color=LoanStatusBinary), size=.3) +
  theme(axis.text.x = element_text(angle=90, vjust=.95, hjust=.95)) +
  facet_wrap(~EmploymentStatus) +
  ggtitle('Loan Amount by Income and Employment')
  #ggsave('visualizations/multivariate/loanamount_income-status-employment.png')

ggplot(aes(x=IncomeRange, y=LoanOriginalAmount), data=subset(df,
    df$IncomeRange!='$0' &  df$IncomeRange!='Not employed' &
      df$IncomeRange!='Not displayed')) +
  geom_jitter(aes(color=LoanStatusBinary), size=.2) +
  theme(axis.text.x = element_text(angle=90, vjust=.95, hjust=.95)) +
  facet_wrap(~ListingCategory..numeric.) +
  ggtitle('Loan Amount by Income and Category')
  #ggsave('visualizations/multivariate/loanamount_income-category-status.png')
```

### Multivariate Analysis

Now that we have analyzed multiple variables together, we are able to see some
clearer patterns, relationships, and trends.

**Loan Amount:** In nearly every scenario analyzed, the loan was directly related
to the loan status, irrespective of any other factor. However, there was one
discover that disproved this. For loans categorized as debt consolidation, higher
loan amounts - even across income brackets - were more likely to be in bad
standing. This defied nearly every other trend across other loan categories.

**APR:** One of the clearest relationships we see is that as credit scores
increase, the borrower's APR decreases, while increasing the likelihood of the
loan being in good standing.

**Homeownership:** While homeowners seem to be borrowing at similar rates and
only slightly higher loan amounts than non-homeowners, non-homeowners were more
likely to have a bad loan status compared to similarly to homeowners.

**Employment:** The safest borrowers were employed, while the most risky were
full-time or retired. Full-time employees were equally risk across the loan
amount and income bracket.

**Income:** As mentioned before, higher income earners borrowed more money at
lower rates. They also borrowed across diverese categories, and were more likely
to have their loan in good standing.

------

## Final Plots and Summary

### Credit and APR
```{r echo=FALSE, Plot_One}
ggplot(aes(x=CreditScoreMean, y=BorrowerAPR), data=subset(df,
    df$IncomeRange!='$0' &  df$IncomeRange!='Not employed' &
      df$IncomeRange!='Not displayed')) +
  geom_jitter(aes(color=LoanStatusBinary), size=.2) +
  theme(axis.text.x = element_text(angle=45, vjust=.95, hjust=.95)) +
  ggtitle('APR by Credit Score and Status')
  #ggsave('visualizations/final/apr_creditscore-status.png')
```

#### Analysis
This visualization demonstates a clear relationship between a borrower's credit
score, their interest rate, and their loan status. In general, borrowers with
higher credit scores borrowed money at lower rates and were more likely to be in
good standing. The inverse is also true; loans with bad standing were likely to
have lower credit scores and higher APR.

### Income and Loan Amount
```{r echo=FALSE, Plot_Two}
ggplot(aes(x=StatedAnnualIncome, y=LoanOriginalAmount), data=subset(df,
    df$IncomeRange!='Not employed' & df$IncomeRange!='Not displayed')) +
  geom_jitter(aes(color=LoanStatusBinary), size=.2) +
  scale_x_continuous(limits = c(1,250000)) +
  theme(axis.text.x = element_text(angle=90, vjust=.5, hjust=.95)) +
  ggtitle('Loan Amount by Income and Status')
  #ggsave('visualizations/final/loanamount_income-status.png')
```

#### Analysis
This plot shows a very clear relationship between a borrower's stated income,
the amount borrowed, and status of the loan. In general, higher income borrowers
borrower more money and are more likely to be in good standing. Conversely,
loans in bad standing are more likely to contain borrowers with a lower income
and smaller loan amounts.

### Loan Categories
```{r echo=FALSE, Plot_Three}
ggplot(aes(x=IncomeRange, y=LoanOriginalAmount), data=subset(df,
    df$IncomeRange!='$0' &  df$IncomeRange!='Not employed' &
      df$IncomeRange!='Not displayed')) +
  geom_jitter(aes(color=LoanStatusBinary), size=.3) +
  theme(axis.text.x = element_text(angle=90, vjust=.95, hjust=.95)) +
  facet_wrap(~ListingCategory..numeric.) +
  ggtitle('Loan Amount by Income and Category')
  #ggsave('visualizations/final/loanamount_income-category-status.png')
```

#### Analysis
Given what we have seen, we know that in general, borrowers with higher credit
scores and higher incomes borrow more money and lower rates, resulting in a
strong direct relationship with good loan status. This visualization shows an
exception to this pattern. For loan's used for debt consolidation, a loan is just
as like to be in a bad status across all income brackets and loan amounts. With
this loan purpose accounting for over 50% of all loans represented in this data
set, this is an important discover to consider.

------

## Reflection

With so much data to analyze, I had to do number bivarate data explorations and
manipulations in order to find patterns and progress forward. Performing a few
multiviarate visualizations made over patterns much easier to spot and identify.