Lesson 3
========================================================

### Pseudo-Facebook User Data

```{r Pseudo-Facebook User Data}
fb_users = read.delim('pseudo_facebook.tsv')
head(fb_users)
names(fb_users)
```

***

### Histogram of Users' Birthdays

```{r Histogram of Users\' Birthdays}
#install.packages('ggplot2')
library(ggplot2)
ggplot(aes(x = fb_users$dob_day), data = fb_users) + geom_histogram(binwidth=1) + scale_x_continuous(breaks=1:31)
```

***
### Faceting

```{r Faceting}
ggplot(aes(x = fb_users$dob_day), data = fb_users) + geom_histogram(binwidth=1) + 
  scale_x_continuous(breaks=1:31) + facet_wrap(~fb_users$dob_month, ncol=3)
```

***

### Friend Count

```{r Friend Count}
ggplot(aes(x=fb_users$friend_count), data=fb_users) + geom_histogram(binwidth = 25) +
  scale_x_continuous(limits=c(0,1000), breaks=seq(0,1000,50))
```

```{r}
summary(fb_users$gender)
```


### Faceting Friend Count
```{r Faceting Friend Count}
# What code would you add to create a facet the histogram by gender?
ggplot(aes(x=friend_count), data=fb_users) +
  geom_histogram() +
  scale_x_continuous(limits=c(0,1000), breaks=seq(0,1000,50)) +
  facet_wrap(~gender)
```

***

### Omitting NA Values

```{r Omitting NA Values}
ggplot(aes(x=friend_count), data=na.omit(fb_users)) +
  geom_histogram() +
  scale_x_continuous(limits=c(0,1000), breaks=seq(0,1000,50)) +
  facet_wrap(~gender)
```

***

### Statistics 'by' Gender

```{r Statistics \'by\' Gender}
table(fb_users$gender)
by(fb_users$friend_count, fb_users$gender, summary)
```

***

### Tenure

```{r Tenure}
ggplot(aes(x=tenure/365), data=fb_users) + 
  geom_histogram(binwidth = .25, color='white') +
  scale_x_continuous(breaks = seq(0,7,1), limits = c(0,7)) +
  xlab('Number of years using Facebook') +
  ylab('Number of users') +
  ggtitle('User Tenure Distribution by Year')
```

***

### User Ages

```{r User Ages}
summary(fb_users$age)
ggplot(aes(x=fb_users$age), data=fb_users) +
  geom_histogram(binwidth = 1, color='white') +
  scale_x_continuous(breaks=seq(0,113,5))
```

***

### Transforming Data
```{r}
summary(fb_users$friend_count)
summary(log10(fb_users$friend_count + 1)) #add 1 to avoid log(0)=infinity
summary(sqrt(fb_users$friend_count))
```

### Graph Results
```{r}
#plot historgrams on one plot
#install.packages("gridExtra")
library(gridExtra)

p1 = ggplot(aes(x=friend_count), data=fb_users) + geom_histogram()
p2 = p1 + scale_x_log10()
p3 = p1 + scale_x_sqrt()

grid.arrange(p1, p2, p3)
```


***

### Add a Scaling Layer
Notes:

```{r Add a Scaling Layer}
scale1 = qplot(x=log10(fb_users$friend_count), data=fb_users)
scale2 = ggplot(aes(x=fb_users$friend_count), data=fb_users) +
  geom_histogram() + scale_x_log10()

grid.arrange(scale1, scale2, ncol=2)
```

***


### Frequency Polygons

```{r Frequency Polygons}
p_friends1 = ggplot(aes(x=friend_count, y=..density..), data=subset(fb_users, !is.na(gender))) +
  geom_freqpoly(aes(color=gender), binwidth=20) +
  scale_x_continuous(limits=c(0,1000))

p_friends2 = ggplot(aes(x=friend_count), data=subset(fb_users, !is.na(gender))) +
  geom_freqpoly(aes(color=gender)) +
  scale_x_continuous(limits=c(0,1000)) +
  scale_x_log10()

p_friends3 = ggplot(aes(x=friend_count, y=..density..), data=subset(fb_users, !is.na(gender))) +
  geom_freqpoly(aes(color=gender)) +
  scale_x_continuous(limits=c(0,1000)) +
  scale_x_log10()

grid.arrange(p_friends1, p_friends2, p_friends3)
```

***

### Likes on the Web
Notes:

```{r Likes on the Web}
ggplot(aes(x=www_likes), data=subset(fb_users, !is.na(gender))) +
  geom_freqpoly(aes(color=gender)) +
  scale_x_log10()

by(fb_users$www_likes, fb_users$gender, sum)
```


***

### Box Plots

### Box Plots, Quartiles, and Friendships
Notes:

```{r Box Plots, Quartiles, and Friendships}
ggplot(aes(x=gender, y=friend_count), data=subset(fb_users, !is.na(gender))) +
  geom_boxplot() +
  coord_cartesian(ylim=c(0,500))

by(fb_users$friend_count, fb_users$gender, summary)
```

#### On average, who initiated more friendships in our sample: men or women?
Response:

```{r Friend Requests by Gender}
names(fb_users)
by(fb_users$friendships_initiated, fb_users$gender, summary)

ggplot(aes(x=gender, y=friendships_initiated), data=subset(fb_users, !is.na(gender))) +
  geom_boxplot() +
  coord_cartesian(ylim=c(0,150))
```

Response:

***

### Getting Logical
Notes:

```{r Getting Logical}
summary(fb_users$mobile_likes > 0)
mobile_check_in = ifelse(fb_users$mobile_likes > 0, 1, 0)
summary(mobile_check_in)['Mean']
```