---
output:
  pdf_document: default
  html_document: default
---
### Investigating Social Network Statistics
#### Muti-Variable Analysis and Plotting


```{r Scatterplots}
library(ggplot2)
df = read.csv('pseudo_facebook.tsv', sep='\t')

head(df)
```

#### Initial Plotting Investigations
```{r Initial investigations}
ggplot(aes(x=age, y=friend_count), data=df) +
  geom_point() +
  ggtitle('Friend Count by Age')
ggsave('visualizations/FriendsbyAge.png')
```


### Overplotting
Notes:

```{r Overplotting}
ggplot(aes(x=age, y=friend_count), data=df) +
  geom_jitter(alpha = .05) + xlim(13,90) +
  ggtitle('Friend Count by Age')
ggsave('visualizations/FriendsbyAge-Enhanced.png')
```

### Coord_trans()
Notes:

```{r Coord_trans()}
ggplot(aes(x=age, y=friend_count), data=df) +
  geom_jitter(alpha = .05) +
  scale_y_sqrt() +
  #geom_point(alpha=.05, position = position_jitter(h=0))
  #coord_trans(y='sqrt') +
  xlim(13,90) +
  ggtitle('Friend Count by Age')
ggsave('visualizations/FriendsbyAge-Scaled.png')
```

```{r Investigate Friendships Initiated by Age}
#names(df)
ggplot(aes(x=age, y=friendships_initiated), data=na.omit(df)) +
  geom_jitter(alpha=.05) +
  scale_y_sqrt() +
  xlim(13,90) +
  facet_wrap(~gender) +
  ggtitle('Friendships by Age from Gender')
ggsave('visualizations/FriendshipsbyAge_Gender.png')
```


### Conditional Means
Notes:

```{r Conditional Means}
#install.packages('dplyr')
library(dplyr)
```

```{r Grouping Dataframes with dplyr}
#Create dataframe by age group
age_groups = group_by(df, age)
age_groups

#create summary dataframe
df.fc_by_age = summarise(age_groups,
                         friend_count_mean=mean(friend_count),
                         friend_count_med=median(friend_count),
                         n=n())
head(df.fc_by_age)
```


Create your plot!

```{r Conditional Means Plot}
ggplot(aes(x=age, y=friend_count_mean), data=df.fc_by_age) +
  geom_line() +
  ggtitle('Mean Friend Count by Age')
ggsave('visualizations/MeanFriendsbyAge.png')
```

***

### Overlaying Summaries with Raw Data
Notes:

```{r Overlaying Summaries with Raw Data}
ggplot(aes(x=age, y=friendships_initiated), data=na.omit(df)) +
  geom_jitter(alpha=.05, color='gray') +
  coord_cartesian(xlim=c(13,90), ylim=c(0,1000)) +
  geom_line(stat='summary', fun.y = mean) +
  geom_line(stat='summary', fun.y = quantile, fun.args=(list(probs=.5)), color='red') +
  geom_line(stat='summary', fun.y = quantile, fun.args=(list(probs=.9)), linetype=2, size=.5) +
  geom_line(stat='summary', fun.y = quantile, fun.args=(list(probs=.1)), linetype=2, size=.5) +
  facet_wrap(~gender) +
  ggtitle('Friendships Initiated by Age by Gender with Statistic Overlays')
ggsave('visualizations/FriendshipsbyAgeGender_Stats.png')
```

***

### Correlation
Notes:

```{r Correlation}
?cor.test
cor.test(df$age, df$friend_count, method=c('pearson'))
```

Look up the documentation for the cor.test function.

What's the correlation between age and friend count? Round to three decimal places.
Response:

***

### Correlation on Subsets
Notes:

```{r Correlation on Subsets}
with(subset(df,age<=70), cor.test(age, friend_count))
```

## Create Scatterplots
Notes:

```{r Likes Received}
sum(df$www_likes_received)
sum(df$likes_received)
ggplot(aes(x=www_likes_received, y=likes_received), data=df) +
  geom_point(alpha=.05) +
  xlim(c(0,quantile(df$www_likes_received,.95))) +
  ylim(c(0,quantile(df$likes_received,.95))) +
  xlab('Desktop Likes Received') +
  ylab('All Likes Received') +
  geom_smooth(method='lm', color='red') +
  ggtitle('Desktop Likes vs Total Likes')
ggsave('visualizations/DesktopLikes_Likes.png')
```

What's the correlation betwen the two variables? Include the top 5% of values for the variable in the calculation and round to 3 decimal places.

```{r Correlation Calcuation}
cor.test(df$www_likes_received, df$likes_received)
```


### More Caution with Correlation
Notes:

```{r More Caution With Correlation}
#install.packages('alr3')
library(alr3)
data(Mitchell)
?Mitchell
names(Mitchell)
```

Create your plot!

```{r Temp vs Month}
ggplot(aes(x=Month, y=Temp), data=Mitchell) +
  geom_point() +
  scale_x_continuous(breaks=seq(0,204,12)) +
  ggtitle('Temperature vs Month (Mitchell Data)')
```

***

### Noisy Scatterplots
What is the actual correlation of the two variables?
(Round to the thousandths place)

```{r Noisy Scatterplots}
cor.test(Mitchell$Month, Mitchell$Temp)
```

***

### Understanding Noise: Age to Age Months
Notes:

```{r Understanding Noise: Age to Age Months}
head(df.fc_by_age)
df$age_with_month = df$age + (1-df$dob_month/12)
head(df)
```

***

### Age with Months Means

```{r Age with Months Means}
#library(dplyr);
age_month_group = group_by(df, age_with_month)
head(age_month_group)

df.fc_by_age_month = summarise(age_month_group,
                               friend_mean_count = mean(friend_count),
                               friend_med_count = median(friend_count),
                               n = n())

df.fc_by_age_month
```

***

### Noise in Conditional Means

```{r Noise in Conditional Means}
p11 = ggplot(aes(x=age_with_month, y=friend_mean_count), data=subset(df.fc_by_age_month, age_with_month<(71))) +
  geom_line() + geom_smooth()

p12 = ggplot(aes(x=age, y=friend_count), data=subset(df, age_with_month<(71))) +
  geom_line(stat='summary', fun.y=mean) + geom_smooth()

p13 = ggplot(aes(x=round(age/5)*5, y=friend_count), data=subset(df, age_with_month<(71))) +
  geom_line(stat='summary', fun.y=mean)

library(gridExtra)
grid.arrange(p11, p12, p13)
g = arrangeGrob(p11, p12, p13)
ggsave('visualizations/FriendsbyAge_comps.png', g)
```